---
title: "CaseStudy01"
author: "Ghizlane Zerouali"
date: "March 15, 2017"
output: 
  html_document:
    keep_md: true
---


# Introduction

The following project involves the reading of the relationship between the revenue monetary that is located in the gross domestic product data (GDP), and the individual income for each country which is found in the educational statistic data. The GDP is a ranking table of different countries made in 2015 that describes the monetary values for each country. The Education Statistics file holds over 4,000 internationally comparable indicators that is last updated on February 2017 and it describe education access, progression, Income, population, and expenditures. 


# Links

* Source Code and Analysis can be found in the GitHub Repository at the following link:
       https://github.com/gzerouali/CaseStudy01
       
* The Gross Domestic Product data for the 190 ranked countries in this data set:
    https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv 
    
* The educational data from this data set:
    https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv  
    

# Libraries 
There are some libraries that are required for this project and primarily intended to clean, tide, merge, and prepare data for analysis.

    * Library (downloader)
    * Library (plyr)
    * Library(ggplot2)
    * Library(reshape2)

# Prepare data for merging
## Downloading the data
The data are downloaded from the following websites:
```{r}
library(downloader)
gdpdata <- download("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv", destfile="gdpdata.csv") 
edstatsdata <- download("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv", destfile="edstatsdata.csv")
```
Reading the data:
```{r}
gdp <- read.csv("gdpData.csv", stringsAsFactors = FALSE, skip = 5, nrows = 190, header = FALSE)
edstats <- read.csv("edstatsdata.csv", stringsAsFactors = F, header = TRUE)
head(gdp,10)
tail(gdp,10)
```

## Cleaning and tidying the data
The gross domestic product data needs to remove regions from it and keep only countries.
```{r}
gdp1 <- gdp[1:190,]
```

Apply the clonames to change the columns title and tide the data, convert GDP values to numeric. Tide the education statistic data set by country code and income group.
```{r}
names(gdp)
colnames(gdp) <- c("CountryCode", "GDP_Rank", "V3", "CountryName", "GDP","V6", "V7", "V8", "V9", "V10")
colnames(gdp)
gdp <- subset(gdp, select = c("CountryCode", "GDP_Rank", "CountryName", "GDP"))
names(gdp)
gdp$GDP_Rank <- as.numeric(gdp$GDP_Rank)
gdp$GDP <- as.numeric(gdp$GDP)
str(gdp)
edstats <- subset(edstats, select = c("CountryCode", "Income.Group"))
names(edstats)

```

Remove entries with no abreviation and remove comma "," to allow coercion.
```{r}
gdp <- gdp[gdp$CountryCode !="", ]
gdp$GDP <- gsub(",", "", gdp$GDP)
```

# The merge
Merge the gross domestic product data set and the education statistic set based on the country shortcode.
```{r}
gdp.edstats <- merge(edstats, gdp, by="CountryCode")
names(gdp.edstats)
```

# Matched ID 
The dimension of the merged data helps to find the number of how many matched countries.
```{r}
dim(gdp.edstats)
sum(is.na(gdp.edstats))
```
After merging the two data sets by country code, the number of 189 present the number of matched IDs in the data frame. 

# Sorting the data
Sort the data frame in descending order by GDP so United States is last.
```{r}
library(plyr)
gdp.edstats <- arrange(gdp.edstats, desc(GDP_Rank))
head(gdp.edstats)
arrange(gdp.edstats, desc(GDP_Rank))[13, ]
```
The 13th country in the resulting data frame is “St. Kitts and Nevis”.

# The average
The average GDP rankings for the "High income: OECD" and "High income: nonOECD" groups.
```{r}
average1 <- mean(subset(gdp.edstats, Income.Group %in% "High income: OECD", select = c(GDP_Rank))$GDP_Rank)
average2 <- mean(subset(gdp.edstats, Income.Group %in% "High income: nonOECD", select = c(GDP_Rank))$GDP_Rank)
average <- c(average1, average2)
average
```

# The plot
The distribution of GDP value for all the countries by income group. 
```{r}
library(ggplot2)
ggplot(gdp.edstats,aes(y=GDP, x=Income.Group))+ labs(title="Income.Group VS GDP")+geom_bar(stat="identity", fill="steelblue")+xlab("Income.Group")+ylab("GDP")+theme( axis.text.x=element_text(angle=90))
```

# Summary
To conclude, the plot of the distribution of GDP value by income group for all the countries shows a strong association between the two variables. The GDP for the lower middle income countries is clearly higher than the other income countries groups. And also the average GDP rankings for the "High income: OECD" is less than the "High income: nonOECD" groups.

# Table
Make a table in depend of income group by deviding the GDP ranking into 5 separate quantile groups, to see How many countries are Lower middle income among the 38 nations with highest GDP.
```{r}
breaks <- quantile(gdp.edstats$GDP_Rank,seq(0, 1, 0.2))
head(cut(gdp.edstats$GDP_Rank, breaks = breaks))
gdp.edstats$quantile <- cut(gdp.edstats$GDP_Rank, breaks=breaks)
table(gdp.edstats$Income.Group, gdp.edstats$quantile)
```

# References

[1] http://www.statmethods.net/management/sorting.html

[2] http://www.statmethods.net/advgraphs/ggplot2.html

[3] http://vita.had.co.nz/papers/tidy-data.pdf

[4] https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html

[5] http://rmarkdown.rstudio.com/articles_intro.html

